---
title: ''
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(nrlR)
library(elo)
library(fuzzyjoin)
library(tidyverse)
library(tidymodels)
library(here)
library(DBI)
library(odbc)
library(httr)
library(jsonlite)
library(validate)
library(janitor)

read_params <- function(){
  
  list(
    season = 2026,
    round_number = 14,
    week_start = 2,
    
    lambda = 0.94, # Degree of recency bias in rolling form variables
    elo_k = 23, # Degree of recency bias in elo model
    
    special_round = "", # "nrl-team-lists-magic-round"
    
    position_minutes_weights = tribble(
      ~position, ~w,
      "Fullback", 1,
      "Wing", 1,
      "Centre", 1,
      "Five-Eighth", 1,
      "Halfback", 1,
      "Prop", 0.6,
      "Hooker", 0.8,
      "Second Row", 0.75,
      "Lock", 0.75,
      "Interchange", 0.4
    ), # weighting stats by minutes played (approximate)
    
    comp = 12999
  )
}

params <- read_params()

# lineup url for scraping
build_lineup_url <- function(params) {
  
  tuesday <- gsub("-", "/", floor_date(Sys.Date(), "week", week_start = params$week_start))
  
  if (params$special_round != ""){
    paste0("https://www.nrl.com/news/", tuesday, "/", params$special_round, "/")
  } else {
    paste0("https://www.nrl.com/news/", tuesday, "/nrl-team-lists-round-", params$round, "/")
  }
}

# upcoming lineups
fetch_upcoming_lineups <- function(lineup_url) {

  current_hour <- as.integer(format(Sys.time(), "%H"))
  
  if (current_hour < 16){
    message("Before 4pm → skipping lineup fetch") # lineups typically get published 4pm AEST
    return(NULL)
  }
    
    nrlR::fetch_lineups(
      url = lineup_url,
      source = "nrl.com",
      type = "team_list"
    )

}
```

# Data Ingestion

## Load Historical Data

```{r}
historical_ladder <- read_csv(here("Data/01_Raw/Ladder Data.csv"))
historical_player <- read_csv(here("Data/01_Raw/Player Data.csv"))
historical_team <- read_csv(here("Data/01_Raw/Team Data.csv"))
player_key <- read_csv(here("Data/01_Raw/Player Key.csv"))
```

## Fetch Previous Round Data

```{r}
PreviousRound_ladder_raw <- nrlR::fetch_ladder_nrl(
  season = params$season, 
  round_number = params$round_number-1
  )

PreviousRound_player_raw <- nrlR::fetch_player_stats(
  season = params$season, 
  round = params$round_number-1,
  league = "nrl", 
  source = "championdata", 
  comp = params$comp
  )

PreviousRound_team_raw <- nrlR::fetch_team_stats_championdata(
  round = params$round_number-1,
  comp = params$comp
  )
```

## Fetch Odds

```{r}
get_nrl_odds <- function(api_key) {
  
  sports_resp <- GET(
    "https://api.the-odds-api.com/v4/sports/",
    query = list(apiKey = api_key)
  )
  
  stop_for_status(sports_resp)
  
  sports <- fromJSON(content(sports_resp, "text", encoding = "UTF-8"))
  
  nrl_sport_df <- sports |> 
    filter(grepl("NRL|Rugby League", title, ignore.case = TRUE) | grepl("rugbyleague", key, ignore.case = TRUE))
  
  if(nrl_sport_df[1,]$active == FALSE){
    stop("NRL is not active: The active column contains FALSE")
  }
  
  odds_resp <- GET(
    "https://api.the-odds-api.com/v4/sports/rugbyleague_nrl/odds/",
    query = list(
      apiKey = api_key,
      regions = "au",
      markets = "h2h",
      oddsFormat = "decimal",
      dateFormat = "iso"
    )
  )
  status_code(odds_resp)
  content(odds_resp, "text", encoding = "UTF-8")
  stop_for_status(odds_resp)
  
  odds_json <- fromJSON(content(odds_resp, "text", encoding = "UTF-8"), flatten = TRUE)
  
  headers(odds_resp)[c("x-requests-remaining","x-requests-used","x-requests-last")]
  
  odds_df <- odds_json |> 
    select(commence_time, home_team, away_team, bookmakers) |> 
    unnest(bookmakers) |> 
    rename(bookmaker_key = key, bookmaker_last_update = last_update) |>
    unnest(markets) |> 
    unnest(outcomes) |> 
    transmute(
      commence_time,
      home_team,
      away_team,
      bookmaker = bookmaker_key,
      market = key, 
      outcome = name,
      odds = price,
      last_update
    ) |> 
    filter(market == "h2h") |> 
    mutate(outcome = case_when(
      outcome == home_team ~ "home_team_odds",
      outcome == away_team ~ "away_team_odds",
      T ~ NA
    )) |> 
    pivot_wider(
      names_from = outcome, values_from = odds
    ) |> 
    mutate(commence_time = as_datetime(commence_time),
           last_update = as_datetime(last_update))
  
  return(odds_df)
}

# Function that caches, limits, and keeps fetch robust
odds_API_fetch <- function() {
  
  api_key <- Sys.getenv("API_KEY")
  
  odds_api <- get_nrl_odds(api_key)
        
  odds_api |> 
    mutate(home_team = str_extract(home_team, "\\w+$"),
           away_team = str_extract(away_team, "\\w+$"),
           date = as_datetime(commence_time))
}
```

```{r}
odds_raw <- odds_API_fetch()
```

## Fetching Upcoming Data

```{r functions}
build_lineup_url <- function(params) {
  
  tuesday <- gsub("-", "/", floor_date(Sys.Date(), "week", week_start = params$week_start))
  
  if (params$special_round != ""){
    paste0("https://www.nrl.com/news/", tuesday, "/", params$special_round, "/")
  } else {
    paste0("https://www.nrl.com/news/", tuesday, "/nrl-team-lists-round-", params$round, "/")
  }
}
```

```{r}
UpcomingRound_ladder_raw <- fetch_ladder_nrl(
  season = params$season, 
  round_number = params$round_number
  )

lineup_url <- build_lineup_url(params)

UpcomingRound_lineups_raw <- nrlR::fetch_lineups(
  url = lineup_url, 
  source = "nrl.com", 
  type = "team_list"
  )

UpcomingRound_fixtures_raw <- nrlR::fetch_results(
  seasons = params$season,
  league = "nrl", 
  source = "rugbyleagueproject"
  ) |> 
  filter(round == params$round_number + 1) # fetching only upcoming round
```

# Cleaning Data

## Cleaning Historical Data

```{r functions}
clean_PreviousRound_ladder <- function(PreviousRound_ladder_raw, params) {
  
  PreviousRound_ladder_raw |> 
    janitor::clean_names() |> # cleans names if format changes
    rename(ladder_points = points) |> # clearer name
    arrange(
      desc(ladder_points), 
      desc(points_diff), 
      desc(points_for)
      ) |> # nrl ladder order
    mutate(
      season = params$season,
      round = params$round_number - 1,
      ladder_position = row_number(), # after arrangin in ladder order, row_number is the position
      team = case_when(
        team == "Wests Tigers" ~ "Tigers",
        team == "Sea Eagles" ~ "Eagles",
        T ~ team)
      ) |> 
    select(colnames(historical_ladder)) # select columns from historical data
  
}

clean_PreviousRound_player <- function(PreviousRound_player_raw, params) {
  
  PreviousRound_player_raw |> 
    mutate(
      season = params$season,
      utc_start = as_datetime(utc_start)
      ) |> 
    janitor::clean_names() |>  # cleans names if format changes
    select(colnames(historical_player)) # select columns from historical data
  
}

clean_PreviousRound_team <- function(PreviousRound_team_raw, params) {
  
  PreviousRound_team_raw |> 
    mutate(
      season = params$season,
      utc_start = as_datetime(utc_start)
      ) |> 
    janitor::clean_names() |>
    select(colnames(historical_team))
  
}
```

```{r}
PreviousRound_ladder_clean <- clean_PreviousRound_ladder(PreviousRound_ladder_raw, params)

PreviousRound_player_clean <- clean_PreviousRound_player(PreviousRound_player_raw, params)

PreviousRound_team_clean <- clean_PreviousRound_team(PreviousRound_team_raw, params)
```

## Cleaning Upcoming Data

### Clean Upcoming Ladder

```{r ladder functions}

clean_UpcomingRound_ladder <- function(UpcomingRound_ladder_raw, historical_ladder) {
  
  UpcomingRound_ladder_raw |> 
    janitor::clean_names() |> # cleans names if format changes
    rename(ladder_points = points) |> # clearer name
    arrange(
      desc(ladder_points), 
      desc(points_diff), 
      desc(points_for)
    ) |> # nrl ladder order
    mutate(
      season = params$season,
      round = params$round_number,
      ladder_position = row_number(), # after arrangin in ladder order, row_number is the position
      team = case_when(
        team == "Wests Tigers" ~ "Tigers",
        team == "Sea Eagles" ~ "Eagles",
        T ~ team)
      ) |> 
    select(colnames(historical_ladder)) # select columns from historical data
}

```

```{r}
UpcomingRound_ladder_clean <- clean_UpcomingRound_ladder(UpcomingRound_ladder_raw, historical_ladder)
```

### Clean Upcoming Fixtures

```{r fixture functions}
map_city_to_team_name <- function(team_name) {
  case_when(
          team_name == "Brisbane" ~ "Brisbane Broncos",
          team_name == "Newcastle" ~ "Newcastle Knights",
          team_name == "Melbourne" ~ "Melbourne Storm",
          team_name == "Cronulla" ~ "Cronulla-Sutherland Sharks",
          team_name == "Manly" ~ "Manly-Warringah Sea Eagles",
          team_name == "North Qld" ~ "North Queensland Cowboys",
          team_name == "South Sydney" ~ "South Sydney Rabbitohs",
          team_name == "Sydney" ~ "Sydney Roosters",
          team_name == "Penrith" ~ "Penrith Panthers",
          team_name == "Parramatta" ~ "Parramatta Eels",
          team_name == "St Geo Illa" ~ "St George-Illawarra Dragons",
          team_name == "Gold Coast" ~ "Gold Coast Titans",
          team_name == "Canberra" ~ "Canberra Raiders",
          team_name == "Canterbury" ~ "Canterbury-Bankstown Bulldogs",
          TRUE ~ team_name)
}

clean_UpcomingRound_fixtures <- function(UpcomingRound_fixtures_raw, params) {
  
  UpcomingRound_fixtures_clean_wide <- UpcomingRound_fixtures_raw |> 
    mutate(across(
        c(home_team, away_team),
        ~map_city_to_team_name(.x)
        )) |> # keeping consistent formatting
    arrange(date) |> 
    mutate(
      match_id = (params$comp*10000L+params$round_number*100L) + row_number(),
      time_clean = str_remove(time, "^[A-Za-z]{3}\\s+"),
      utc_start = ymd_hm(paste(date, time)) 
    ) |> # keeping consistent formatting
    select(match_id, round, home = home_team, away = away_team, utc_start, season)
  
    UpcomingRound_fixtures_clean_wide |> 
    pivot_longer(
      home:away,
      values_to = "team_name",
      names_to = "team_location"
    ) |> 
    mutate(team = str_extract(team_name, "\\w+$")) |> 
    select(team, match_id, round, season, team_location, utc_start)
  
}
```

```{r}
UpcomingRound_fixtures_clean_long <- clean_UpcomingRound_fixtures(UpcomingRound_fixtures_raw, params)
```

### Clean Upcoming Lineup

```{r lineup functions}
clean_player_firstname <- function(firstname) {
  
  case_when(
      firstname == "KL" ~ "Kayal",
      firstname == "Sebastian" ~ "Seb",
      firstname == "Leka" ~ "Selumiela",
      T ~ firstname
      )
}
clean_player_surname <- function(surname) {
  
  case_when(
      surname == "Vaimauga" ~ "Sifakula",
      T ~ surname
      )
}

build_UpcomingRound_lineup_NaiveMatch <- function(UpcomingRound_lineups_raw, player_key) {
  
  UpcomingRound_lineups_raw |> 
    rename(
      firstname = first_name, 
      surname = last_name,
      position = role) |> 
    mutate(
      team = ifelse(team == "Wests Tigers", "Tigers", team),
      team = ifelse(team == "Sea Eagles", "Eagles", team),
      position = case_when(
        position == "Winger" ~ "Wing",
        position == "2nd" ~ "Second Row",
        T ~ position),
      firstname = clean_player_firstname(firstname), # Cleaning historical culprits
      surname = clean_player_surname(surname) # Cleaning historical culprits
      ) |> # keeping consistent formatting
    filter(position != "Reserve") |> # reserves rarely get on, so remove for now
    left_join(player_key, by = c("firstname", "surname")) # naive join with player key
  
}


pull_UpcomingRound_lineup_UnMatched <- function(UpcomingRound_lineup_NaiveMatch) {
  
  UpcomingRound_lineup_NaiveMatch |> 
    filter(is.na(player_id)) |> 
    select(-player_id)
  
}


build_UpcomingRound_lineup_FuzzyMatch <- function(UpcomingRound_lineup_UnMatched, player_key) {
  
  UpcomingRound_lineup_UnMatched |> 
    stringdist_left_join( 
      player_key,
      by = c("firstname", "surname"),
      method = "jw",
      max_dist = 0.22 # Found that this is the sweet spot historically, reproducible because we manually update player_key now
      ) |> 
    rename(firstname = firstname.x, surname = surname.x) |> 
    select(-firstname.y, -surname.y)
  
}


pull_UpcomingRound_lineup_NewPlayers <- function(UpcomingRound_lineup_FuzzyMatch) {
  
  UpcomingRound_lineup_FuzzyMatch |> 
      filter(is.na(player_id)) |> 
      select(firstname, surname, player_id)
  
}


update_player_key <- function(UpcomingRound_lineup_NewPlayers, player_key) {
  
  player_key |> 
    bind_rows(
      UpcomingRound_lineup_NewPlayers
    ) |> 
    arrange(player_id) |> 
    mutate(
      player_id = accumulate(player_id, ~ ifelse(is.na(.y), .x + 1, .y))
    )
  
}


pull_UpComingRound_lineup_NewPlayers_key <- function(UpcomingRound_lineup_NewPlayers, updated_player_key) {
  
  UpcomingRound_lineup_NewPlayers |> 
    select(-player_id) |> 
    left_join(
      updated_player_key, 
      by = c("firstname", "surname")
      )
  
}


combine_UpcomingRound_lineup_matches <- function(UpcomingRound_lineup_NaiveMatch, UpcomingRound_lineup_FuzzyMatch, UpComingRound_lineup_NewPlayers_key) {
  
  UpcomingRound_lineup_NaiveMatch |> 
    drop_na(player_id) |> 
    bind_rows(
      UpcomingRound_lineup_FuzzyMatch |> 
        left_join(
          UpComingRound_lineup_NewPlayers_key, 
          by = c("firstname", "surname")
          ) |> 
        rename(player_id = player_id.x) |> 
        mutate(player_id = ifelse(is.na(player_id), player_id.y, player_id)) |> 
        select(-player_id.y)
      ) |> 
    select(player_id, firstname, surname, team, position)
  
}


join_UpcomingRound_fixtures_to_lineup <- function(UpcomingRound_lineup_combined, UpcomingRound_fixtures_clean_long) {
  
  UpcomingRound_lineup_combined |> 
    left_join(
      UpcomingRound_fixtures_clean_long,
      by = c("team")
    ) |> 
    rename(team_name = team)
  
}
```

```{r}
UpcomingRound_lineup_NaiveMatch <- build_UpcomingRound_lineup_NaiveMatch(UpcomingRound_lineups_raw, player_key)

UpcomingRound_lineup_UnMatched <- pull_UpcomingRound_lineup_UnMatched(UpcomingRound_lineup_NaiveMatch)

UpcomingRound_lineup_FuzzyMatch <- build_UpcomingRound_lineup_FuzzyMatch(UpcomingRound_lineup_UnMatched, player_key)

UpcomingRound_lineup_NewPlayers <- pull_UpcomingRound_lineup_NewPlayers(UpcomingRound_lineup_FuzzyMatch)

updated_player_key <- update_player_key(UpcomingRound_lineup_NewPlayers, player_key)

UpComingRound_lineup_NewPlayers_key <- pull_UpComingRound_lineup_NewPlayers_key(UpcomingRound_lineup_NewPlayers, updated_player_key)

UpcomingRound_lineup_combined <- combine_UpcomingRound_lineup_matches(UpcomingRound_lineup_NaiveMatch, UpcomingRound_lineup_FuzzyMatch, UpComingRound_lineup_NewPlayers_key)

UpcomingRound_lineup_clean <- join_UpcomingRound_fixtures_to_lineup(UpcomingRound_lineup_combined, UpcomingRound_fixtures_clean_long)
```

# Update Historical Data

```{r functions}
update_historical_ladder <- function(historical_ladder, PreviousRound_ladder_clean) {
  
  bind_rows(
    historical_ladder,
    PreviousRound_ladder_clean
    ) |> 
    distinct(team, season, round, .keep_all = TRUE) # removing duplicates
  
}

update_historical_player <- function(historical_player, PreviousRound_player_clean) {
  
  bind_rows(
    historical_player,
    PreviousRound_player_clean
    ) |> 
    distinct(player_id, match_id, .keep_all = TRUE) # removing duplicates
  
}

update_historical_team <- function(historical_team, PreviousRound_team_clean) {
  
  bind_rows(
    historical_team,
    PreviousRound_team_clean
    ) |> 
    distinct(team_name, match_id, .keep_all = TRUE) # removing duplicates
  
}

update_player_key <- function(updated_player_key) {
  
  updated_player_key |> 
    distinct(player_id, .keep_all = TRUE) # removing duplicates
    
}
```

```{r}
historical_ladder_updated <- update_historical_ladder(historical_ladder, PreviousRound_ladder_clean)

historical_player_updated <- update_historical_player(historical_player, PreviousRound_player_clean)

historical_team_updated <- update_historical_team(historical_team, PreviousRound_team_clean)

player_key_updated <- update_player_key(updated_player_key)
```

```{r}#
write_csv(historical_ladder_updated, here("Data/01_Raw/Ladder Data.csv"))
write_csv(historical_player_updated, here("Data/01_Raw/Player Data.csv"))
write_csv(historical_team_updated, here("Data/01_Raw/Team Data.csv"))
write_csv(player_key_updated, here("Data/01_Raw/Player Key.csv"))
```

```{r}
historical_team_updated <- historical_team
historical_player_updated <- historical_player
historical_ladder_updated <- historical_ladder
```

# Feature Engineering

```{r feature calc functions}
compute_elo <- function(df, k) {
  
  elo.run(
    result ~ team_home + team_away,
    data = df,
    k = k
  ) |> 
    as_tibble() |> 
  bind_cols(df |> select(utc_start)) |> # select date from elo input
  arrange(desc(utc_start)) |> 
  select(team_home = team.A, team_away = team.B, elo_home = elo.A, elo_away = elo.B, utc_start) |> 
  pivot_longer(
    cols = team_home:elo_away,
    names_to = c(".value", "team_location"),
    names_sep = "_"
  ) |> 
  arrange(utc_start) |> 
  group_by(team) |> 
  mutate(elo = lag(elo),
         elo = replace_na(elo, 1500)) |> 
    ungroup()
}

# input vector x, and degree of weighting for rolling forms lambda. Get lagged rolling forms output.
rolling_lambda <- function(x, lambda) {
  lag(purrr::accumulate(x, ~lambda*.x + (1-lambda)*.y), 1)
}
```

## Team Level Features

```{r team functions}

build_TeamContext_features <- function(historical_team_updated, historical_ladder_updated, UpcomingRound_fixtures_clean_long) {
  
  historical_team_updated |> 
  bind_rows(UpcomingRound_fixtures_clean_long |> rename(team_name = team)) |> 
  group_by(competition_id, team_name) |> 
  arrange(team_name, utc_start) |> 
  mutate(
    days_rest = as.double(utc_start - lag(utc_start)),
    days_rest = ifelse(days_rest > 50, NA, days_rest),
    team_name = str_extract(team_name, "\\w+$"),
    season_stage = factor(case_when(
      round < 12 ~ "Early",
      round >= 12 & round <= 19 ~ "Mid",
      round > 19 ~ "Late"
      ), 
      levels = c("Early", "Mid", "Late"))
    ) |> 
  ungroup() |> 
  left_join(
    historical_ladder_updated |> 
      bind_rows(UpcomingRound_ladder_clean) |> 
      rename(team_name = team), 
    by = c("team_name", "season", "round")
    ) |> 
  select(-comp) |> 
  rename(
    ladder_points_for = points_for, 
    ladder_points_against = points_against, 
    ladder_points_diff = points_diff
    )
  
}

build_elo_input <- function(TeamContext_features) {
  
  TeamContext_features |> 
    select(match_id, utc_start, team_name, team_location, score) |> 
    rename(team = team_name) |> 
    pivot_wider(
      id_cols = c(match_id, utc_start),
      names_from = team_location,
      values_from = c(team, score)
    ) |> 
    mutate(
      result = case_when(
      score_home > score_away ~ 1,
      score_home < score_away ~ 0,
      score_home == score_away ~ 0.5,
      T ~ 0)
      ) |> 
    select(match_id, utc_start, team_home, team_away, score_home, score_away, everything()) |> 
    arrange(utc_start)
  
}

build_elo_features <- function(elo_input, params) {
  
  elo_model <- compute_elo(elo_input, k = params$elo_k)

  recent_elo_model <- compute_elo(elo_input, k = 10) |> 
  rename(shortelo = elo)

  long_elo_model <- compute_elo(elo_input, k = 40) |> 
    rename(longelo = elo)
  
  elo_input |> 
    select(utc_start, match_id, team_home, team_away) |>
    pivot_longer(
      cols = team_home:team_away,
      names_to = c(".value", "team_location"),
      names_sep = "_"
    ) |> 
    left_join(elo_model) |> 
    left_join(recent_elo_model) |> 
    left_join(long_elo_model)
}

combine_TeamLevel_features <- function(TeamContext_features, elo_features, params) {
  
  TeamContext_features |> 
    left_join(
      elo_features |> 
        select(match_id, team, elo, shortelo, longelo) |> 
        rename(team_name = team), 
      by = c("match_id", "team_name")
      ) |> 
    mutate(result = score) |> 
    arrange(utc_start) |> 
    group_by(team_name) |> 
    mutate(across(
      score:goal_line_dropouts,
      ~ rolling_lambda(., params$lambda) # uses rolling forms, lagged so that future stats are not leaked
      )) |> 
    mutate(across(
      ladder_points:ladder_position, # ensures future ladder is not leaked
      ~ lag(.x, n = 1)
    )) |> 
    ungroup()
  
}
```

```{r}
TeamContext_features <- build_TeamContext_features(historical_team_updated, historical_ladder_updated, UpcomingRound_fixtures_clean_long)

elo_input <- build_elo_input(TeamContext_features)

elo_features <- build_elo_features(elo_input, params)

TeamLevel_features <- combine_TeamLevel_features(TeamContext_features, elo_features, params)
```

## Player Level Features

```{r player functions}
build_player_statistics <- function(historical_player_updated, UpcomingRound_lineup_clean, params) {
  
  historical_player_updated |> 
  bind_rows(UpcomingRound_lineup_clean) |> 
  unite("name", firstname:surname, sep = "_") |> # combine into one variable
  mutate(
    possessions = ifelse(possessions == 0, NA, possessions), # helps with division (per possession), immaterial to model as well
    goals_per_kick = (conversions+penalty_goals) / (conversion_attempts+penalty_goal_attempts)
    ) |> 
  select(
    -conversions, 
    -conversion_attempts, 
    -penalty_goals, 
    -penalty_goal_attempts,
    -runs_hitup,
    -runs_normal, 
    -runs_hitup_metres, 
    -runs_normal_metres, 
    -tackleds, 
    -handling_errors, 
    -post_contact_metres
    ) |> # lots of noisy stats that are similar to other stats
  left_join(
    params$position_minutes_weights, # estimate on minutes played in game based on position (not in data)
    by = "position"
    ) |> 
  mutate(
    m_per_run=run_metres/runs,
    tb_per_run=tackle_breaks/runs,
    lb_per_run=line_breaks/runs,
    offs_per_run=offloads/runs, # Attacking Runs
    
    ta_per_touch=try_assists/possessions,
    lba_per_touch=line_break_assists/possessions,
    kick_per_touch=kicks_general_play/possessions,
    pass_per_touch=passes/possessions,
    km_per_kick=ifelse(kicks_general_play ==0, NA, kick_metres/kicks_general_play), # Attacking Play-making
    
    tries_per_game=tries,
    err_per_touch=errors/possessions,
    points_per_game=points, # Attacking Overall
    
    tackles_time=tackles/w,
    saves_per_game=try_saves, # Defensive Work
    
    miss_per_t=missed_tackles/tackles,
    pen_per_t=penalties_conceded/tackles,
    ineff_per_t=tackles_ineffective/tackles # Defensive Discipline
  ) |> 
  select(-(points:penalties_conceded)) |> # We have scaled versions of these
  group_by(player_id) |> 
  arrange(player_id, utc_start) |> 
  select(
    utc_start, 
    name, 
    player_id, 
    team_name, 
    position, 
    match_id, 
    competition_id, 
    season, 
    round,
    team_location, 
    w,
    everything()
    ) |> 
  mutate(across(
    goals_per_kick:ineff_per_t,
    ~ rolling_lambda(., params$lambda) # uses rolling forms, lagged so that future stats are not leaked
    )) |> 
  ungroup()
  
}

combine_player_features <- function(player_statistics) {
  
  player_statistics |> 
    group_by(match_id, team_name) |> 
    summarise(
      utc_start = first(utc_start),
      comp_id = first(competition_id),
      season = first(season),
      round = first(round),
      team_location = first(team_location),
      across(goals_per_kick:ineff_per_t, ~ mean(.x, na.rm = TRUE)) # Averaging statistics across team, may change to weighted averages later
      ) |> 
    ungroup() |> 
    mutate(team_name = str_extract(team_name, "\\w+$"))
  
}
```

```{r}
player_statistics <- build_player_statistics(historical_player_updated, UpcomingRound_lineup_clean, params)

PlayerLevel_features <- combine_player_features(player_statistics)
```

## Combine Features

```{r features functions}
build_features_data <- function(PlayerLevel_features, TeamLevel_features) {
  
  PlayerLevel_features |> 
    left_join(
      TeamLevel_features |> 
        select(-utc_start, -team_location, -competition_id, -round, -season), 
      by = c("match_id", "team_name")) |> 
    select(result, match_id, utc_start, round, season_stage, everything()) |> 
    pivot_wider(
      id_cols = c(match_id, utc_start, round, season_stage),
      names_from = team_location,
      values_from = c(result, team_name, goals_per_kick:longelo),
      names_sep = "_"
      ) |> 
    mutate(
      result = factor(case_when(
        result_home > result_away ~ "H",
        result_away > result_home ~ "A",
        T ~ "A"))
        ) |> # making target variable (binary classification)
    rename(date = utc_start) |> # clearer name
    rename(
      home_team = team_name_home, 
      away_team = team_name_away,
      away_result = result_away, 
      home_result = result_home
      ) |> # cleaning names and also avoiding below function
      mutate(
      across(ends_with("_home"),
           ~ .-get(str_replace(cur_column(), "_home$", "_away")),
           .names = "{str_remove(.col,'_home$')}_diff"
           )
      ) |> # This function takes the difference between _home and _away variables, creating _diff variables.
    mutate(
      prhome = 1 / (1 + 10^( (elo_away - elo_home) / 400) ),
      shortprhome = 1 / (1 + 10^( (shortelo_away - shortelo_home) / 400) ),
      longprhome = 1 / (1 + 10^( (longelo_away - longelo_home) / 400) )
      ) |> # pr of winning based on elo ratings
    mutate(across(
      goals_per_kick_away:ladder_position_diff,
      ~ replace_na(.x, 0)
      )) |> # ok to impute this statistics like this, NAs are typically zeroes or didn't actually get the stat
    mutate(across(
      where(is.numeric),
      ~ ifelse(is.finite(.), ., NA_real_)
      )) # figure out cause of this, only 1 or 2 in a few columns
  
}
```

```{r}
features_data <- build_features_data(PlayerLevel_features, TeamLevel_features)
```

# Update Features Data

```{r}
write_csv(features_data, here("Data/02_Features/feature_engineered_df.csv"))
```

# Fit Model and Predict

## Data Split

```{r}
cutoff <- floor_date(Sys.Date(), unit = "weeks", week_start = 2)

train_data <- features_data |>  filter(date < cutoff)
pred_data  <- features_data |>  filter(date >= cutoff)
```

## Model Parameters

```{r}
model_parameters <- readRDS(here("Outputs/model_parameters.rds"))
```

## Model specification

```{r}
define_XgBoost_model_specification <- function(model_parameters) {
  boost_tree(
    trees = 1000,
    tree_depth = model_parameters$tree_depth,
    min_n = model_parameters$min_n,
    loss_reduction = model_parameters$loss_reduction,
    sample_size = model_parameters$sample_size,
    mtry = model_parameters$mtry,
    learn_rate = model_parameters$learn_rate
  ) |> 
    set_engine(
      "xgboost",
      eval_metric = "logloss",
      nthread = 4
      ) |> 
    set_mode("classification")
}

model_specification <- define_XgBoost_model_specification(model_parameters)
```

## Data Preprocessing Recipe

```{r}

define_XgBoost_Data_PreprocessingRecipe <- function(train_data) {
  
  recipe(result ~ ., data = train_data) |> 
    update_role(match_id, new_role = "ID") |> 
    step_rm(
      home_result, away_result, # target variable proxies
      home_team, away_team, # Noisy, 17 factors
      date, round) |> # Using season stage for this
    step_zv(all_predictors()) |> # Remove predictors that are constant "no predictive value"
    step_novel(all_nominal_predictors()) |> # assigns unseen factor levels "new"
    step_dummy(all_nominal_predictors()) # dummies factor variables
  
}

data_recipe <- define_XgBoost_Data_PreprocessingRecipe(train_data)
```

## Modelling workflow

```{r}
modelling_workflow <- workflow() |> 
  add_model(model_specification) |> 
  add_recipe(data_recipe)
```

## Fitting Model

```{r}
set.seed(234)

model_fit <- fit(modelling_workflow, data = train_data)

upcoming_predictions <- model_fit |> 
  predict(new_data = pred_data, type = "prob") |> 
  bind_cols(pred_data) |> 
  select(
    date,
    match_id,
    home_team,
    away_team,
    home_prediction = .pred_H,
    away_prediction = .pred_A
    ) |> 
  mutate(prediction = ifelse(home_prediction > away_prediction, home_team, away_team))
```

## Model Metadata

```{r}
model_filename <- paste0("Models/", "model_metadata_", params$season, "_round", params$round_number, ".rds")

model_metadata <- list(
  model = model_specification$engine,
  parameters = model_parameters,
  metric = "logloss",
  trained_on = Sys.Date(),
  feature_version = "v1",
  notes = paste0("season ",params$season, " round ", params$round_number, " predictions")
)

write_rds(model_metadata, here(model_filename))
```

# Create Outputs

## Load Outputs Files

```{r}
historical_predictions_log <- read_csv(here("Data/03_Outputs/prediction_log.csv"))

historical_odds_log <- read_csv(here("Data/03_Outputs/odds_log.csv"))

historical_bets_log <- read_csv(here("Data/03_Outputs/bets_log.csv"))
```

## Get New Data

```{r new logs functions}
pull_new_odds_log_data <- function(upcoming_predictions, odds_raw, historical_odds_log) {
  
  upcoming_predictions |> 
    right_join(
      odds_raw,
      by = c("home_team", "away_team")
    ) |> 
    select(colnames(historical_odds_log)) |> 
    drop_na(match_id)
  
}

pull_new_bets_log_data <- function(upcoming_predictions, odds_raw, historical_bets_log) {
  
  upcoming_predictions |> 
    right_join(
      odds_raw,
      by = c("home_team", "away_team")
    ) |> 
    group_by(match_id) |> 
    mutate(
      best_home_odds = max(home_team_odds),
      best_away_odds = max(away_team_odds),
      home_ev = home_prediction * best_home_odds - 1,
      away_ev = away_prediction * best_away_odds - 1,
      ev_bet = ifelse(home_ev > away_ev, home_team, away_team),
      odd_diff = ifelse(home_ev >= away_ev, best_home_odds - home_team_odds, best_away_odds - away_team_odds)
      ) |> 
    arrange(odd_diff) |> 
    select(colnames(historical_bets_log)) |> 
    ungroup() |> 
    drop_na(match_id)
  
}
```

```{r}
new_odds_log_data <- pull_new_odds_log_data(upcoming_predictions, odds_raw, historical_odds_log)

new_bets_log <- pull_new_bets_log_data(upcoming_predictions, odds_raw, historical_bets_log)
```

## Update Outputs

```{r update logs functions}
update_predictions_log <- function(historical_predictions_data, upcoming_predictions) {
  
  bind_rows(
    historical_predictions_log,
    upcoming_predictions
    ) |> 
    distinct(date, away_team, home_team, .keep_all = TRUE) # removing duplicates
  
}

update_odds_log <- function(historical_odds_log, new_odds_log_data) {
  
  bind_rows(
    historical_odds_log,
    new_odds_log_data) |> 
    distinct(match_id, bookmaker, .keep_all = TRUE)
  
}

update_bets_log <- function(historical_bets_log, new_bets_log) {
  
  bind_rows(
    historical_bets_log,
    new_bets_log) |> 
    distinct(match_id, .keep_all = TRUE) # removing duplicates
  
}
```

```{r}
predictions_log_updated <- update_predictions_log(historical_predictions_data, upcoming_predictions)
odds_log_updated <- update_odds_log(historical_odds_log, new_odds_log_data)
bets_log_updated <- update_bets_log(historical_bets_log, new_bets_log)
```

```{r}
write_csv(predictions_log_updated, here("Data/03_Outputs/prediction_log.csv"))
write_csv(odds_log_updated, here("Data/03_Outputs/odds_log.csv"))
write_csv(bets_log_updated, here("Data/03_Outputs/bets_log.csv"))
```



























