---
title: 
output: html_document
date: "2026-05-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(nrlR)
library(elo)
library(fuzzyjoin)
library(tidyverse)
library(tidymodels)
library(here)
library(DBI)
library(odbc)
library(httr)
library(jsonlite)
library(validate)
library(janitor)

read_params <- function(){
  
  list(
    season = 2026,
    round_number = 13,
    week_start = 2,
    
    lambda = 0.94, # Degree of recency bias in rolling form variables
    elo_k = 23, # Degree of recency bias in elo model
    
    special_round = "", # "nrl-team-lists-magic-round"
    
    position_minutes_weights = tribble(
      ~position, ~w,
      "Fullback", 1,
      "Wing", 1,
      "Centre", 1,
      "Five-Eighth", 1,
      "Halfback", 1,
      "Prop", 0.6,
      "Hooker", 0.8,
      "Second Row", 0.75,
      "Lock", 0.75,
      "Interchange", 0.4
    ), # weighting stats by minutes played (approximate)
    
    comp = 12999
  )
}

params <- read_params()

#fetch_cd_comps() |> 
#  filter(str_detect(name, "NRL Finals") | str_detect(name, "NRL Premiership")) |> 
#  slice_tail(n = -4)

# lineup url for scraping
build_lineup_url <- function(params) {
  
  tuesday <- gsub("-", "/", floor_date(Sys.Date(), "week", week_start = params$week_start))
  
  if (params$special_round != ""){
    paste0("https://www.nrl.com/news/", tuesday, "/", params$special_round, "/")
  } else {
    paste0("https://www.nrl.com/news/", tuesday, "/nrl-team-lists-round-", params$round, "/")
  }
}

# upcoming lineups
fetch_upcoming_lineups <- function(lineup_url) {

  current_hour <- as.integer(format(Sys.time(), "%H"))
  
  if (current_hour < 16){
    message("Before 4pm → skipping lineup fetch") # lineups typically get published 4pm AEST
    return(NULL)
  }
    
    nrlR::fetch_lineups(
      url = lineup_url,
      source = "nrl.com",
      type = "team_list"
    )

}
```

# Data Ingestion

## Load Historical Data

```{r}
historical_ladder <- read_csv(here("Data/01_Raw/Ladder Data.csv"))
historical_player <- read_csv(here("Data/01_Raw/Player Data.csv"))
historical_team <- read_csv(here("Data/01_Raw/Team Data.csv"))
player_key <- read_csv(here("Data/01_Raw/Player Key.csv"))
```


## Fetch Previous Round Data

```{r}
raw_PreviousRound_ladder <- nrlR::fetch_ladder_nrl(
  season = params$season, 
  round_number = params$round_number-1
  )

raw_PreviousRound_player <- nrlR::fetch_player_stats(
  season = params$season, 
  round = params$round_number-1,
  league = "nrl", 
  source = "championdata", 
  comp = params$comp
  )

raw_PreviousRound_team <- nrlR::fetch_team_stats_championdata(
  round = params$round_number-1,
  comp = params$comp
  )
```


## Fetch Odds

```{r}
get_nrl_odds <- function(api_key) {
  
  sports_resp <- GET(
    "https://api.the-odds-api.com/v4/sports/",
    query = list(apiKey = api_key)
  )
  
  stop_for_status(sports_resp)
  
  sports <- fromJSON(content(sports_resp, "text", encoding = "UTF-8"))
  
  nrl_sport_df <- sports |> 
    filter(grepl("NRL|Rugby League", title, ignore.case = TRUE) | grepl("rugbyleague", key, ignore.case = TRUE))
  
  if(nrl_sport_df[1,]$active == FALSE){
    stop("NRL is not active: The active column contains FALSE")
  }
  
  odds_resp <- GET(
    "https://api.the-odds-api.com/v4/sports/rugbyleague_nrl/odds/",
    query = list(
      apiKey = api_key,
      regions = "au",
      markets = "h2h",
      oddsFormat = "decimal",
      dateFormat = "iso"
    )
  )
  status_code(odds_resp)
  content(odds_resp, "text", encoding = "UTF-8")
  stop_for_status(odds_resp)
  
  odds_json <- fromJSON(content(odds_resp, "text", encoding = "UTF-8"), flatten = TRUE)
  
  headers(odds_resp)[c("x-requests-remaining","x-requests-used","x-requests-last")]
  
  odds_df <- odds_json |> 
    select(commence_time, home_team, away_team, bookmakers) |> 
    unnest(bookmakers) |> 
    rename(bookmaker_key = key, bookmaker_last_update = last_update) |>
    unnest(markets) |> 
    unnest(outcomes) |> 
    transmute(
      commence_time,
      home_team,
      away_team,
      bookmaker = bookmaker_key,
      market = key, 
      outcome = name,
      odds = price,
      last_update
    ) |> 
    filter(market == "h2h") |> 
    mutate(outcome = case_when(
      outcome == home_team ~ "home_team_odds",
      outcome == away_team ~ "away_team_odds",
      T ~ NA
    )) |> 
    pivot_wider(
      names_from = outcome, values_from = odds
    ) |> 
    mutate(commence_time = as_datetime(commence_time),
           last_update = as_datetime(last_update))
  
  return(odds_df)
}

# Function that caches, limits, and keeps fetch robust
odds_API_fetch <- function() {
  
  api_key <- readRDS(here("Config/api_key.rds")) # fix later
  
  odds_api <- get_nrl_odds(api_key)
        
  odds_api |> 
    mutate(home_team = str_extract(home_team, "\\w+$"),
           away_team = str_extract(away_team, "\\w+$"),
           date = as.POSIXct(commence_time))
}
```

```{r}
odds_api <- get_nrl_odds(api_key)
    
odds_raw <- odds_api |> 
      mutate(home_team = str_extract(home_team, "\\w+$"),
             away_team = str_extract(away_team, "\\w+$"))
```

# Cleaning
## Clean Previous Round Data

```{r}
clean_PreviousRound_ladder <- raw_PreviousRound_ladder |> 
  janitor::clean_names() |> # cleans names if format changes
  rename(ladder_points = points) |> # clearer name
  arrange(
    desc(ladder_points), 
    desc(points_diff), 
    desc(points_for)
  ) |> # nrl ladder order
  mutate(
    season = params$season,
    round = params$round_number - 1,
    ladder_position = row_number(), # after arrangin in ladder order, row_number is the position
    team = case_when(
      team == "Wests Tigers" ~ "Tigers",
      team == "Sea Eagles" ~ "Eagles",
      T ~ team)
    ) |> 
  select(colnames(historical_ladder)) # select columns from historical data

clean_PreviousRound_player <- raw_PreviousRound_player |> 
  mutate(
    season = params$season,
    utc_start = as_datetime(utc_start)
    ) |> 
  janitor::clean_names() |>  # cleans names if format changes
  select(colnames(historical_player)) # select columns from historical data

clean_PreviousRound_team <- raw_PreviousRound_team |> 
  mutate(
    season = params$season,
    utc_start = as_datetime(utc_start)
    ) |> 
  janitor::clean_names() |>
  select(colnames(historical_team))
```


# Update
## Update Historical Data

```{r}
updated_historical_ladder <- bind_rows(
  historical_ladder,
  clean_PreviousRound_ladder
  ) |> 
  group_by(team, season, round) |> 
  slice_tail(n = 1) |> 
  ungroup() # removing duplicates

updated_historical_player <- bind_rows(
  historical_player,
  clean_PreviousRound_player
  ) |> 
  group_by(player_id, match_id) |> 
  slice_tail(n = 1) |> 
  ungroup() # removing duplicates

updated_historical_team <- bind_rows(
  historical_team,
  clean_PreviousRound_team
  ) |> 
  group_by(team_name, match_id) |> 
  slice_tail(n = 1) |> 
  ungroup() # removing duplicates

write_csv(updated_historical_ladder, here("Data/01_Raw/Ladder Data.csv"))
write_csv(updated_historical_player, here("Data/01_Raw/Player Data.csv"))
write_csv(updated_historical_team, here("Data/01_Raw/Team Data.csv"))
```


# Upcoming Data
## Fetch Raw Upcoming Data

```{r}
raw_UpcomingRound_fixtures <- nrlR::fetch_results(
  seasons = params$season,
  league = "nrl", 
  source = "rugbyleagueproject"
  ) |> 
  filter(round == params$round_number + 1) # fetching only upcoming round

lineup_url <- build_lineup_url(params)

raw_UpcomingRound_lineups <- nrlR::fetch_lineups(
  url = lineup_url, 
  source = "nrl.com", 
  type = "team_list"
  )

raw_UpcomingRound_ladder <- fetch_ladder_nrl(
  season = params$season, 
  round_number = params$round_number
  )
```


## Clean Upcoming Data

```{r}
map_city_to_team_name <- function(team_name) {
  case_when(
          team_name == "Brisbane" ~ "Brisbane Broncos",
          team_name == "Newcastle" ~ "Newcastle Knights",
          team_name == "Melbourne" ~ "Melbourne Storm",
          team_name == "Cronulla" ~ "Cronulla-Sutherland Sharks",
          team_name == "Manly" ~ "Manly-Warringah Sea Eagles",
          team_name == "North Qld" ~ "North Queensland Cowboys",
          team_name == "South Sydney" ~ "South Sydney Rabbitohs",
          team_name == "Sydney" ~ "Sydney Roosters",
          team_name == "Penrith" ~ "Penrith Panthers",
          team_name == "Parramatta" ~ "Parramatta Eels",
          team_name == "St Geo Illa" ~ "St George-Illawarra Dragons",
          team_name == "Gold Coast" ~ "Gold Coast Titans",
          team_name == "Canberra" ~ "Canberra Raiders",
          team_name == "Canterbury" ~ "Canterbury-Bankstown Bulldogs",
          TRUE ~ team_name)
}

clean_UpcomingRound_fixtures_wide <- raw_UpcomingRound_fixtures |> 
    mutate(across(
        c(home_team, away_team),
        ~map_city_to_team_name(.x)
        )) |> # keeping consistent formatting
  arrange(date) |> 
  mutate(
    match_id = (params$comp*10000L+params$round_number*100L) + row_number(),
    time_clean = str_remove(time, "^[A-Za-z]{3}\\s+"),
    utc_start = ymd_hm(paste(date, time)) 
  ) |> # keeping consistent formatting
  select(match_id, round, home = home_team, away = away_team, utc_start, season)

clean_UpcomingRound_fixtures_long <- clean_UpcomingRound_fixtures_wide |> 
  pivot_longer(
    home:away,
    values_to = "team_name",
    names_to = "team_location"
  ) |> 
  mutate(team = str_extract(team_name, "\\w+$")) |> 
  select(team, match_id, round, season, team_location, utc_start)
```


```{r}
clean_player_firstname <- function(firstname) {
  
  case_when(
      firstname == "KL" ~ "Kayal",
      firstname == "Sebastian" ~ "Seb",
      firstname == "Leka" ~ "Selumiela",
      T ~ firstname
      )
}
clean_player_surname <- function(surname) {
  
  case_when(
      surname == "Vaimauga" ~ "Sifakula",
      T ~ surname
      )
}
```

```{r}
raw_UpcomingRound_lineup_NaiveMatch <- raw_UpcomingRound_lineups |> 
  rename(
    firstname = first_name, 
    surname = last_name,
    position = role) |> 
  mutate(
    team = ifelse(team == "Wests Tigers", "Tigers", team),
    team = ifelse(team == "Sea Eagles", "Eagles", team),
    position = case_when(
      position == "Winger" ~ "Wing",
      position == "2nd" ~ "Second Row",
      T ~ position),
    firstname = clean_player_firstname(firstname), # Cleaning historical culprits
    surname = clean_player_surname(surname) # Cleaning historical culprits
    ) |> # keeping consistent formatting
  filter(position != "Reserve") |> # reserves rarely get on, so remove for now
  left_join(player_key, by = c("firstname", "surname")) # naive join with player key


raw_UpcomingRound_lineup_UnMatched <- raw_UpcomingRound_lineup_NaiveMatch |> 
  filter(is.na(player_id)) |> 
  select(-player_id)


raw_UpcomingRound_lineup_FuzzyMatch <- stringdist_left_join( 
  raw_UpcomingRound_lineup_UnMatched,
  player_key,
  by = c("firstname", "surname"),
  method = "jw",
  max_dist = 0.22
) |> 
  rename(firstname = firstname.x, surname = surname.x) |> 
  select(-firstname.y, -surname.y)


raw_UpcomingRound_lineup_NewPlayers <- raw_UpcomingRound_lineup_FuzzyMatch |> 
      filter(is.na(player_id)) |> 
      select(firstname, surname, player_id)


updated_player_key <- player_key |> 
  bind_rows(
    raw_UpcomingRound_lineup_NewPlayers
  ) |> 
  arrange(player_id) |> 
  mutate(
    player_id = accumulate(player_id, ~ ifelse(is.na(.y), .x + 1, .y))
  )
write_csv(updated_player_key, here("Data/01_Raw/Player Key.csv"))


NewPlayer_key <- raw_UpcomingRound_lineup_NewPlayers |> 
  select(-player_id) |> 
  left_join(updated_player_key, by = c("firstname", "surname"))


clean_UpcomingRound_lineup <- raw_UpcomingRound_lineup_NaiveMatch |> 
  drop_na(player_id) |> 
  bind_rows(
    raw_UpcomingRound_lineup_FuzzyMatch |> 
      left_join(NewPlayer_key, by = c("firstname", "surname")) |> 
      rename(player_id = player_id.x) |> 
      mutate(player_id = ifelse(is.na(player_id), player_id.y, player_id)) |> 
      select(-player_id.y)
    ) |> 
  select(player_id, firstname, surname, team, position) |> 
  left_join(
    clean_UpcomingRound_fixtures_long,
    by = c("team")
  ) |> 
  rename(team_name = team)
```


```{r}
clean_UpcomingRound_ladder <- raw_UpcomingRound_ladder |> 
  janitor::clean_names() |> # cleans names if format changes
  rename(ladder_points = points) |> # clearer name
  arrange(
    desc(ladder_points), 
    desc(points_diff), 
    desc(points_for)
  ) |> # nrl ladder order
  mutate(
    season = params$season,
    round = params$round_number,
    ladder_position = row_number(), # after arrangin in ladder order, row_number is the position
    team = case_when(
      team == "Wests Tigers" ~ "Tigers",
      team == "Sea Eagles" ~ "Eagles",
      T ~ team)
    ) |> 
  select(colnames(historical_ladder)) # select columns from historical data
```

# Engineer Features

```{r}
compute_elo <- function(df, k) {
  
  elo.run(
    result ~ team_home + team_away,
    data = df,
    k = k
  ) |> 
    as_tibble() |> 
  bind_cols(df |> select(utc_start)) |> # select date from elo input
  arrange(desc(utc_start)) |> 
  select(team_home = team.A, team_away = team.B, elo_home = elo.A, elo_away = elo.B, utc_start) |> 
  pivot_longer(
    cols = team_home:elo_away,
    names_to = c(".value", "team_location"),
    names_sep = "_"
  ) |> 
  arrange(utc_start) |> 
  group_by(team) |> 
  mutate(elo = lag(elo),
         elo = replace_na(elo, 1500)) |> 
    ungroup()
}

# input vector x, and degree of weighting for rolling forms lambda. Get lagged rolling forms output.
rolling_lambda <- function(x, lambda) {
  lag(purrr::accumulate(x, ~lambda*.x + (1-lambda)*.y), 1)
}

```

```{r}
team_rest_ladder_features <- updated_historical_team |> 
  bind_rows(clean_UpcomingRound_fixtures_long |> rename(team_name = team)) |> 
  group_by(competition_id, team_name) |> 
  arrange(team_name, utc_start) |> 
  mutate(
    days_rest = as.double(utc_start - lag(utc_start)),
    days_rest = ifelse(days_rest > 50, NA, days_rest),
    team = str_extract(team_name, "\\w+$")
    ) |> 
  ungroup() |> 
  left_join(
    updated_historical_ladder |> 
      bind_rows(clean_UpcomingRound_ladder), 
    by = c("team", "season", "round")
    ) |> 
  select(-team, -comp) |> 
  rename(
    ladder_points_for = points_for, 
    ladder_points_against = points_against, 
    ladder_points_diff = points_diff
    )
```

```{r}
# Elos

elo_input <- team_rest_ladder_features |> 
  select(match_id, utc_start, team_name, team_location, score) |> 
  rename(team = team_name) |> 
  pivot_wider(
    id_cols = c(match_id, utc_start),
    names_from = team_location,
    values_from = c(team, score)
  ) |> 
  mutate(
    result = case_when(
    score_home > score_away ~ 1,
    score_home < score_away ~ 0,
    score_home == score_away ~ 0.5,
    T ~ 0
  )) |> 
  arrange(utc_start)

elo_model <- compute_elo(elo_input, k = params$elo_k)

recent_elo_model <- compute_elo(elo_input, k = 10) |> 
  rename(shortelo = elo)

long_elo_model <- compute_elo(elo_input, k = 40) |> 
    rename(longelo = elo)

team_elo_features <- elo_input |> 
  select(utc_start, match_id, team_home, team_away) |>
  pivot_longer(
    cols = team_home:team_away,
    names_to = c(".value", "team_location"),
    names_sep = "_"
  ) |> 
  left_join(elo_model) |> 
  left_join(recent_elo_model) |> 
  left_join(long_elo_model)
```

```{r}
team_features <- team_rest_ladder_features |> 
  left_join(
    team_elo_features |> 
      select(match_id, team, elo, shortelo, longelo) |> 
      rename(team_name = team), 
    by = c("match_id", "team_name")
    ) |> 
  mutate(result = score) |> 
  arrange(utc_start) |> 
  group_by(team_name) |> 
  mutate(across(
    score:goal_line_dropouts,
    ~ rolling_lambda(., params$lambda) # uses rolling forms, lagged so that future stats are not leaked
    )) |> 
  mutate(across(
    ladder_points:ladder_position, # ensures future ladder is not leaked
    ~ lag(.x, n = 1)
  )) |> 
  ungroup() |> 
  mutate(
    season_stage = factor(case_when(
      round < 12 ~ "Early",
      round >= 12 & round <= 19 ~ "Mid",
      round > 19 ~ "Late"
      ), 
      levels = c("Early", "Mid", "Late"))
    )
```

```{r}

player_statistics <- updated_historical_player |> 
  bind_rows(clean_UpcomingRound_lineup) |> 
  unite("name", firstname:surname, sep = "_") |> 
  mutate(
    possessions = ifelse(possessions == 0, NA, possessions), 
    goals_per_kick = (conversions+penalty_goals) / (conversion_attempts+penalty_goal_attempts)
    ) |> 
  select(
    -conversions, 
    -conversion_attempts, 
    -penalty_goals, 
    -penalty_goal_attempts,
    -runs_hitup,
    -runs_normal, 
    -runs_hitup_metres, 
    -runs_normal_metres, 
    -tackleds, 
    -handling_errors, 
    -post_contact_metres
    ) |> 
  left_join(
    params$position_minutes_weights, 
    by = "position"
    ) |> 
  mutate(
    m_per_run=run_metres/runs,
    tb_per_run=tackle_breaks/runs,
    lb_per_run=line_breaks/runs,
    offs_per_run=offloads/runs, # Attacking Runs
    
    ta_per_touch=try_assists/possessions,
    lba_per_touch=line_break_assists/possessions,
    kick_per_touch=kicks_general_play/possessions,
    pass_per_touch=passes/possessions,
    km_per_kick=ifelse(kicks_general_play ==0, NA, kick_metres/kicks_general_play), # Attacking Play-making
    
    tries_per_game=tries,
    err_per_touch=errors/possessions,
    points_per_game=points, # Attacking Overall
    
    tackles_time=tackles/w,
    saves_per_game=try_saves, # Defensive Work
    
    miss_per_t=missed_tackles/tackles,
    pen_per_t=penalties_conceded/tackles,
    ineff_per_t=tackles_ineffective/tackles # Defensive Discipline
  ) |> 
  select(-(points:penalties_conceded)) |> 
  group_by(player_id) |> 
  arrange(player_id, utc_start) |> 
  select(
    utc_start, 
    name, 
    player_id, 
    team_name, 
    position, 
    match_id, 
    competition_id, 
    season, 
    round,
    team_location, 
    w,
    everything()
    ) |> 
  mutate(across(
    goals_per_kick:ineff_per_t,
    ~ rolling_lambda(., params$lambda) # uses rolling forms, lagged so that future stats are not leaked
    )) |> 
  ungroup()
```

```{r}
player_features <- player_statistics |> 
  group_by(match_id, team_name) |> 
  summarise(
    utc_start = first(utc_start),
    comp_id = first(competition_id),
    season = first(season),
    round = first(round),
    team_location = first(team_location),
    across(goals_per_kick:ineff_per_t, ~ mean(.x, na.rm = TRUE)) # Averaging statistics across team, may change to weighted averages later
    ) |> 
  ungroup()
```

```{r}
features_data <- player_features |> 
  left_join(
    team_features |> 
      select(-utc_start, -team_location, -competition_id, -round, -season), 
    by = c("match_id", "team_name")) |> 
  select(result, everything()) |> 
  pivot_wider(
    id_cols = c(match_id, utc_start, round, season_stage),
    names_from = team_location,
    values_from = c(result, team_name, goals_per_kick:longelo),
    names_sep = "_"
  ) |> 
  mutate(
    result = factor(case_when(
      result_home > result_away ~ "H",
      result_away > result_home ~ "A",
      T ~ "A"))
    ) |> # making target variable (binary classification)
  rename(date = utc_start) |> # clearer name
  rename(
    home_team = team_name_home, 
    away_team = team_name_away,
    away_result = result_away, 
    home_result = result_home
    ) |> # cleaning names and also avoiding below function
    mutate(
    across(ends_with("_home"),
           ~ .-get(str_replace(cur_column(), "_home$", "_away")),
           .names = "{str_remove(.col,'_home$')}_diff"
           )
    ) |> # This function takes the difference between _home and _away variables, creating _diff variables.
  mutate(
    prhome = 1 / (1 + 10^( (elo_away - elo_home) / 400) ),
    shortprhome = 1 / (1 + 10^( (shortelo_away - shortelo_home) / 400) ),
    longprhome = 1 / (1 + 10^( (longelo_away - longelo_home) / 400) )
  ) |> # pr of winning based on elo ratings
  mutate(across(
    goals_per_kick_away:ladder_position_diff,
    ~ replace_na(.x, 0)
    )) |> # ok to impute this statistics like this, NAs are typically zeroes or didn't actually get the stat
  mutate(across(
    where(is.numeric),
    ~ ifelse(is.finite(.), ., NA_real_)
    )) # figure out cause of this, only 1 or 2 in a few columns

write_csv(features_data, here("Data/02_Features/feature_engineered_df.csv"))
```


# Fit Model and Predict

## Data Split

```{r}
cutoff <- floor_date(Sys.Date(), unit = "weeks", week_start = 2)

train_data <- features_data |>  filter(date < cutoff)
pred_data  <- features_data |>  filter(date >= cutoff)
```

## Model Parameters

```{r}
model_parameters <- readRDS(here("Outputs/model_parameters.rds"))
```

## Model specification

```{r}
define_XgBoost_model_specification <- function(model_parameters) {
  boost_tree(
    trees = 1000,
    tree_depth = model_parameters$tree_depth,
    min_n = model_parameters$min_n,
    loss_reduction = model_parameters$loss_reduction,
    sample_size = model_parameters$sample_size,
    mtry = model_parameters$mtry,
    learn_rate = model_parameters$learn_rate
  ) |> 
    set_engine(
      "xgboost",
      nthread = 4,
      objective = "binary:logistic",
      eval_metric = "logloss"
      ) |> 
    set_mode("classification")
}


model_specification <- define_XgBoost_model_specification(model_params)
```

## Data Preprocessing Recipe

```{r}
data_recipe <- recipe(result ~ ., data = train_data) |> 
  update_role(match_id, new_role = "ID") |> 
  step_rm(
    home_result, away_result, # target variable proxies
    home_team, away_team, # Noisy, 17 factors
    date, round) |> # Using season stage for this
  step_zv(all_predictors()) |> # Remove predictors that are constant "no predictive value"
  step_novel(all_nominal_predictors()) |> # assigns unseen factor levels "new"
  step_dummy(all_nominal_predictors()) # dummies factor variables
```

## Modelling workflow

```{r}
modelling_workflow <- workflow() |> 
  add_model(model_specification) |> 
  add_recipe(data_recipe)
```

## Fitting Model

```{r}
set.seed(234)

model_fit <- fit(modelling_workflow, data = train_data)

upcoming_predictions <- model_fit |> 
  predict(new_data = pred_data, type = "prob") |> 
  bind_cols(pred_data) |> 
  select(
    date,
    match_id,
    home_team,
    away_team,
    home_prediction = .pred_H,
    away_prediction = .pred_A
    ) |> 
  mutate(prediction = ifelse(home_prediction > away_prediction, home_team, away_team))
```

## Model Metadata

```{r}
model_filename <- paste0("Models/", "model_metadata_", params$season, "_round", params$round_number, ".rds")

model_metadata <- list(
  model = model_specification$engine,
  parameters = model_parameters,
  metric = "logloss",
  trained_on = Sys.Date(),
  feature_version = "v1",
  notes = paste0("season ",params$season, " round ", params$round_number, " predictions")
)

write_rds(model_metadata, here(model_filename))
```


# Create Outputs

## Predictions Log

```{r}
historical_predictions_data <- read_csv(here("Data/03_Outputs/prediction_log.csv"))

updated_predictions <- bind_rows(
  historical_predictions_data,
  upcoming_predictions
  ) |> 
  group_by(date, away_team, home_team) |> 
  slice_head(n = 1) |> # Removes duplicate records
  ungroup() 

write_csv(predictions, here("Outputs/prediction_log.csv"))
```

## Odds Log

```{r}
historical_odds_log <- read_csv(here("Data/03_Outputs/odds_log.csv")) |> 
  mutate(
    date = as_datetime(date),
    last_update = as_datetime(last_update)
    )

new_odds_log_data <- upcoming_predictions |> 
  right_join(
      odds_raw,
      by = c("home_team", "away_team")
  ) |> 
  select(colnames(historical_odds_log)) |> 
  drop_na(match_id)

updated_odds_log <- bind_rows(
  historical_odds_log,
  new_odds_log_data) |> 
  group_by(match_id, bookmaker) |> 
  slice_tail(n = 1) |> # Keeps updated version instead? Improve this later.
  ungroup()

write_csv(updated_odds_log, here("Data/03_Outputs/odds_log.csv"))
```

## Bets Log

```{r}
historical_bets_log <- read_csv(here("Data/03_Outputs/bets_log.csv")) |> 
  mutate(date = as_datetime(date)) |> 
  drop_na(match_id)

new_bets_log <- upcoming_predictions |> 
  right_join(
      odds_raw,
      by = c("home_team", "away_team")
  ) |> 
  group_by(match_id) |> 
  mutate(
    best_home_odds = max(home_team_odds),
    best_away_odds = max(away_team_odds),
    home_ev = home_prediction * best_home_odds - 1,
    away_ev = away_prediction * best_away_odds - 1,
    ev_bet = ifelse(home_ev > away_ev, home_team, away_team),
    odd_diff = ifelse(home_ev >= away_ev, best_home_odds - home_team_odds, best_away_odds - away_team_odds)
    ) |> 
  arrange(odd_diff) |> 
  select(colnames(historical_bets_log)) |> 
  ungroup() |> 
  drop_na(match_id)

updated_bets_log <- bind_rows(
  historical_bets_log,
  new_bets_log) |> 
  group_by(match_id) |> 
  slice_tail(n = 1) |> # Keeps updated version instead? Improve this later.
  ungroup()

write_csv(updated_bets_log, here("Data/03_Outputs/bets_log.csv"))
```



























































